---
title: "Lab 9"
author: "Kajal Patel"
date: "3/25/2020"
output: html_document
---
# Making a graph from the daily reports
```{r}
library(tidyverse)
library(lubridate)
```
# Plotting data from 3/11/2020
```{r}
report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country.Region = "Country/Region", Province.State = "Province/State")
```
# Checking table properties
```{r}
head(report_03_11_2020)
```
```{r}
str(report_03_11_2020)
```
# Plot of confirmed cases in the US on 3/11/2020
```{r}
report_03_11_2020 %>% 
  filter (Country.Region == "US") %>% 
  ggplot(aes(x = Confirmed,  y = reorder(Province.State, Confirmed))) + 
    geom_point() +
    ggtitle("Confirmed cases for each US State") +
    ylab("Country/Region") +
    xlab("Confirmed Cases")

```
# Summarizing the data being graphed
```{r}
report_03_11_2020 %>% 
  group_by(Country.Region) %>% 
  summarise(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths))
```
# Plotting countries with the most deaths as of 3/11/2020
```{r}
report_03_11_2020 %>% 
  group_by(Country.Region) %>% 
  summarise(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:20) %>% 
  ggplot(aes(x = Deaths,  y = reorder(Country.Region, Deaths))) + 
    geom_bar(stat = 'identity') +
    ggtitle("The 20 countries with the most reported COV19-related deaths") +
    ylab("Country/Region") +
    xlab("Deaths")
```
# Working with the time series data
# Loading the most recent time series data for confirmed cases
```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province.State = "Province/State", Country.Region = "Country/Region")
```
# Checking the table
```{r}
head(time_series_confirmed)
```
# Converting the data to a long format
```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province.State, Country.Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") %>% 
               group_by(Country.Region,Date) %>% 
               summarise(Confirmed = sum(Confirmed))
```
# Convert date to data format
```{r}
time_series_confirmed_long$Date <- mdy(time_series_confirmed_long$Date)
```
# Looking at the format of the data frame
```{r}
head(time_series_confirmed_long)
```
# A graph of the confirmed COVID-19 cases in the US
```{r}
time_series_confirmed_long %>% 
  filter (Country.Region == "US") %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("US Confirmed COVID-19 Cases")
```
# Plotting more countries
```{r}
time_series_confirmed_long %>% 
  filter (Country.Region %in% c("China","Japan", "Korea, South",
                                "Italy","Spain", "US")) %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("Confirmed COVID-19 Cases") +
    facet_wrap(~Country.Region, ncol=2, scales="free_y")
```
# Having several countries on the same graph
```{r}
time_series_confirmed_long %>% 
  filter (Country.Region %in% c("China","France","Italy", 
                                "Korea, South", "US")) %>% 
    ggplot(aes(x = Date,  y = Confirmed, color = Country.Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("Confirmed COVID-19 Cases")
```
# Visualizing data on maps
# Loading the additinal libraries
```{r}
library(maps)
library(viridis)
```
# Getting the world map
```{r}
world <- map_data("world")
```
# Cutoffs based on the number of cases
```{r}
mybreaks <- c(1, 20, 100, 1000, 50000)
```
# Plotting the data from 3/11/2020
```{r}
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_confirmed, aes(x=Long, y=Lat, size=`2/25/20`, color=`2/25/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )
```
# Getting data to longer format and summarizing
```{r}
time_series_confirmed_long2 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
    rename(Province.State = "Province/State", Country.Region = "Country/Region") %>%
    pivot_longer(-c(Province.State, Country.Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))

```
# Viewing the table
```{r}
head(time_series_confirmed_long2)
```
# Plotting the data
```{r}
static <- ggplot(subset(time_series_confirmed_long2,Date %in% seq(min(Date),max(Date),7)),
            aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
            borders("world", colour = NA, fill = "grey90") +
            theme_bw() +
            geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
            labs(title = 'COVID-19 spread',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,nrow=3)
static
```
# Exercises
# A new graph from the daily reports
```{r}
report_03_24_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-24-2020.csv")) %>%
  rename(Country.Region = "Country_Region", Province.State = "Province_State")
```
# Checking the table
```{r}
head(report_03_24_2020)
```
```{r}
str(report_03_24_2020)
```
#Plotting the confirmed cases in China as of 3/24/2020
```{r}
report_03_24_2020 %>% 
  filter (Country.Region == "China") %>% 
  ggplot(aes(x = Confirmed,  y = reorder(Province.State, Confirmed))) + 
    geom_point() +
    ggtitle("Confirmed Cases in China") +
    ylab("Province") +
    xlab("Confirmed Cases")
```
# Plotting the countries with the most reported COVID-19 recoveries
```{r}
report_03_24_2020 %>% 
  group_by(Country.Region) %>% 
  summarise(Recovered = sum(Recovered)) %>% 
  arrange(desc(Recovered)) %>% 
  slice(1:10) %>% 
  ggplot(aes(x = Recovered,  y = reorder(Country.Region, Recovered))) + 
    geom_bar(stat = 'identity') +
    ggtitle("The 20 countries with the most reported COVID-19 recoveries") +
    ylab("Country") +
    xlab("Recovered")
```
# New graphs with the time series data
```{r}
time_series_deaths <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province.State = "Province/State", Country.Region = "Country/Region")
```
# Checking the table
```{r}
head(time_series_deaths)
```
# Converting it to a long format
```{r}
time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province.State, Country.Region, Lat, Long),
                            names_to = "Date", values_to = "deaths") %>% 
               group_by(Country.Region,Date) %>% 
               summarise(deaths = sum(deaths))
```
```{r}
time_series_deaths_long$Date <- mdy(time_series_deaths_long$Date)

```
# Looking at the format of the data frame
```{r}
head(time_series_deaths_long)
```
# A graph of the number of COVID-19 related deaths in India
```{r}
time_series_deaths_long %>% 
  filter (Country.Region == "India") %>% 
    ggplot(aes(x = Date,  y = deaths)) + 
    geom_point(color=" dark green") +
    geom_line(color="dark green") +
    ggtitle("COVID-19 Related Deaths in India")
```
# Comparing the number of COVID-19 related deaths in different countries
```{r}
time_series_deaths_long %>% 
  filter (Country.Region %in% c("India","Canada","Norway", 
                                "Brazil", "Egypt")) %>% 
    ggplot(aes(x = Date,  y = deaths, color = Country.Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Related Deaths")
```
# Creating new geographical maps
# Getting a world map and making cutoffs based on the number of cases
```{r}
world <- map_data("world")
mybreaks <- c(1, 500, 1000, 3000, 5000)
```
#Plotting the data
```{r}
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="light grey", alpha=0.3) +
  geom_point(data=time_series_deaths, aes(x=Long, y=Lat, size=`3/24/20`, color=`3/24/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="deaths", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-100", "101-500", "501-1000", "1,001-3000", "3,001+")) +
  # scale_alpha_continuous(name="deaths", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="deaths", trans="log",breaks=mybreaks, labels = c("1-100", "101-500", "501-1000", "1,001-3,000", "3,001+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme (
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
)
```
# Second Geographical Map
```{r}
time_series_deaths_long2 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
    rename(Province.State = "Province/State", Country.Region = "Country/Region") %>%
    pivot_longer(-c(Province.State, Country.Region, Lat, Long),
    names_to = "Date", values_to = "deaths") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(deaths = ifelse(sum(deaths)>0,
        sum(deaths),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))
```
# Viewing the table
```{r}
head(time_series_deaths_long)
```
#Plotting the data
```{r}
static <- ggplot(subset(time_series_deaths_long2,Date %in% seq(min(Date),max(Date),21)),
            aes(x = Long, y = Lat, size = deaths/100)) +
            borders("world", colour = NA, fill = "grey90") +
            theme_bw() +
            geom_point(shape = 21, color='sky blue', fill='sky blue', alpha = 0.5) +
            labs(title = 'COVID-19 Deaths',x = '', y = '',
                 size="deaths (x10))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,nrow=3)
static
```























